The error when inferring phylogenies with incipient species by a Birth-Death model

### 2018-04-17 article.tex
Fixed flaw

From an incipient species tree, we sample a species tree. To do
so, from each species a sub-species must be chosen to represent
the good species as a whole. To clarify, it may be that an
incipient species branched of from a good species. Both of these
subspecies are recognized as the good species, and both can be 
picked as a representative. Here we use three sampling scenario's,
in which we pick the youngest (most recent incipient speciation), 
oldest or random subspecies. Note that only in few 
cases (when speciation completion rate is low) the incipient species
chosen will have an effect on the branch length distribution of species 
tree. The simplest scenario requires a species, with one good and
one incipient lineage, after which either lineages gives 
rise to a new good species, before the first incipient lineage
completes speciation itself.

@@

There will be two data sets produced by this research.
The first data set will be a general balanced data set to chart
the effect of the biological parameters on the nLTT statistic
distribution. In this data set, incipient species are sampled 
randomly to represent a good species.
The second data set charts the effect of sampling
subspecies and only uses PBD trees in which this sampling
has an effect. For each of these trees, we sample both youngest and
oldest subspecies. We predict that these two most extreme sampling methods
result in the most pronounced differences.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l l}
    \hline
    Parameter             & Description & Values \\
    \hline
    \hline
    $b = b_g = b_i$       & Speciation initiation rate & 0.1, 0.5, 1.0 \\
    $\lambda$             & Speciation completion rate & 0.1, 0.3, 1.0, $\infty$ \\
    $\mu = \mu_g = \mu_i$ & Extinction rate & 0.0, 0.1, 0.2, 0.4 \\
    \hline
    $t_c$                 & Crown age & 15 \\
    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
    $M$                   & Sampling method & Random \\
    $r$                   & Mutation rate & $\frac{1000}{15}$ \\
    $l_a$                 & DNA alignment length & $15K$ \\
    $f_i$                 & MCMC sampling interval & 1K or more \\
    $R_i$                 & RNG seed incipient tree & 1 to 40K \\
    $R_a$                 & RNG seed alignment simulation & $R_i$ \\
    $R_b$                 & RNG seed BEAST2 & $R_i$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the 12 simulation parameters. Above the horizontal line is 
    the biological parameter set.
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    $n$ & Description \\
    \hline
    \hline
    $12$   & simulation parameters, see table \ref{table:parameters} \\
    $1000$ & nLTT statistic values \\
    $1$ & [NOTE: keep this in?] expected duration of speciation (from Etienne \& Rosindell 2012) \\
    $1$ & [NOTE: keep this in?] number of taxa in incipient species tree (also: number of subspecies) \\
    $1$ & [NOTE: keep this in?] number of taxa in species tree (also: number of species) \\
    $1$ & [NOTE: keep this in?] mean nLTT statistic \\
    $1$ & [NOTE: keep this in?] median nLTT statistic \\
    $1$ & [NOTE: keep this in?] lower bound of 95\% HPD value of nLTT statistic \\
    $1$ & [NOTE: keep this in?] upper bound of 95\% HPD value of nLTT statistic \\
    $11$ & ESSes of all parameters estimated by BEAST2 (see specs below) \\
    \hline
  \end{tabular}
  \caption{
    Specification of the data sets. $n$ denotes the number
    of columns a certain item will occupy, resulting in a table of 
    1030 columns and 40K rows.
  }
  \label{table:specs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

From an incipient species tree, we sample a species tree. To do
so, from each species a sub-species must be chosen to represent
the good species as a whole. To clarify, it may be that an
incipient species branched of from a good species. Both of these
subspecies are recognized as the good species, and both can be 
picked as a representative. Here we use three sampling scenario's,
in which we pick the most recent common ancestor (MRCA), 
most distant common ancestor (MDCA) or random subspecies. Note that only in few 
cases (when speciation completion rate is low) the incipient species
chosen will have an effect on the branch length distribution of species 
tree. The simplest scenario requires a species, with one good and
one incipient lineage, after which either lineages gives 
rise to a new good species, before the first incipient lineage
completes speciation itself.

@@

There will be two data sets produced by this research.
The first data set will be a general balanced data set to chart
the effect of the biological parameters on the nLTT statistic
distribution. In this data set, incipient species are sampled 
randomly to represent a good species.
The second data set charts the effect of sampling
subspecies and only uses PBD trees in which this sampling
has an effect. For each of these trees, we sample both MRCA and
MDCA subspecies. We predict that these two most extreme sampling methods
result in the most pronounced differences.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l l}
    \hline
    Parameter             & Description & Values \\
    \hline
    \hline
    $b = b_g = b_i$       & Speciation initiation rate & 0.1, 0.5, 1.0 \\
    $\lambda$             & Speciation completion rate & 0.1, 0.3, 1.0, $\infty$ \\
    $\mu = \mu_g = \mu_i$ & Extinction rate & 0.0, 0.1, 0.2, 0.4 \\
    \hline
    $t_c$                 & Crown age & 15 \\
    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
    $M$                   & Sampling method & MRCA, MDCA or random \\
    $r$                   & Mutation rate & $\frac{1000}{15}$ \\
    $l_a$                 & DNA alignment length & $15K$ \\
    $f_i$                 & MCMC sampling interval & 1K or more \\
    $R_i$                 & RNG seed incipient tree & 1 to 40K \\
    $R_a$                 & RNG seed alignment simulation & $R_i$ \\
    $R_b$                 & RNG seed BEAST2 & $R_i$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the 12 simulation parameters. Above the horizontal line is 
    the biological parameter set. Sampling method $M$ is random for the general
    data set. For the data set exploring the effect of sampling, MRCA is
    used for odd values of $R_i$, and MDCA is used for even values of $R_i$.
    $R_i$ is 1 for the first simulation, 2 for the next, etcetera.
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    $n$ & Description \\
    \hline
    \hline
    $12$   & simulation parameters, see table \ref{table:parameters} \\
    $1000$ & nLTT statistic values \\
    $11$ & ESSes of all parameters estimated by BEAST2 (see specs below) \\
    \hline
  \end{tabular}
  \caption{
    Specification of the data sets. $n$ denotes the number
    of columns a certain item will occupy, resulting in a table of 
    1023 columns and 40K rows.
  }
  \label{table:specs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
